---
title: "setting_up_api_keys"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{setting_up_api_keys}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(usdc)
```

## What is an API key?

Like many other R packages that are API clients, `usdc` lets you set up and use an API key.
An API key is a string that you communicate to the server (see [Wikipedia Page](https://en.wikipedia.org/wiki/Application_programming_interface_key)). The key is used by the server to identify you. This helps the server to manage your access, for example, giving access to specific services.


## Why use one?

Many APIs have a limit in the number of queries you can make in a given amount of time. Using an API key, because it identifies you, lets you go over this limit so the server knows it is not under a [Denial of Service (DoS) attack](https://en.wikipedia.org/wiki/Denial-of-service_attack).

## Registering for an API key

### Bitquery
You can register a free API key at <https://graphql.bitquery.io/ide>
You have to provide your name and a valid email address to register for an API key.

The Bitquery API has a limit of 100,000 free calls per month on the Free plan tier currently

## How to set up your API key in `usdc`

Now that you have registered for an API key, you can use it in `usdc`. There are several ways to set up your API key in `usdc`.


### Using `Sys.setenv()` during an interactive session

One way to set up your API key without using the `key` argument is to use **environment variables**. Environment variables are system or user-defined variables that are used to store information during your R session.

`usdc` searches for different environment variables for each available API service:
- `BITQUERY_KEY` : for Bitquery
- `THEGRAPH_KEY` : for The Graph's USDC subgraph
- `BLOCKCHAIR_KEY` : for Blockchair

```
Sys.getenv("BITQUERY_KEY")
```

If none has been defined it should return `""`. You can then set the environment variable using the function `Sys.setenv()`:


```r
Sys.setenv(BITQUERY_KEY = "BQY8IPItkNCP0iK8BqvmhSWV1pxpFqrB")
Sys.getenv("BITQUERY_KEY")
#> [1] "BQY8IPItkNCP0iK8BqvmhSWV1pxpFqrB"
```

It is set up for the rest of your R session and `usdc` will automatically use it when you call the different functions. You can also use the `check_key()` function that should give the same result if your key is well set-up:


```r
library("usdc")
check_key()
#> [1] "BQY8IPItkNCP0iK8BqvmhSWV1pxpFqrB"
```


### Using the dedicated function `*_auth()`

`usdc` provides the function `rr_auth()` that creates a well named environmental variable that can be used in the rest of your session:


```r
bitquery_auth("BQY8IPItkNCP0iK8BqvmhSWV1pxpFqrB")
check_key()
#> [1] "BQY8IPItkNCP0iK8BqvmhSWV1pxpFqrB"
```

This works so that you don't have to specify the name of the environmental variable by hand. Under the hood `bitquery_auth()` uses the same mechanism as explained in the above mentioned section.


### Setting up your API key in an `.Rprofile` file

Every time R starts it looks for `.Rprofile` files in different locations:

- `R_HOME` the directory in which R is installed,
- `HOME` the user's home directory,
- R's current working directory.

R only loads one `.Rprofile` file per session and thus an `.Rprofile` file at the project-level overrides files in other locations.

The `.Rprofile` file is an R script that is launch each time R starts. Put it at the root of your project and type the following:

```
BQY8IPItkNCP0iK8BqvmhSWV1pxpFqrB = "BQY8IPItkNCP0iK8BqvmhSWV1pxpFqrB"
```

You can then reload your session and check that `usdc` managed to get your key by using the `check_key()` function:


```r
check_key()
#> [1] "BQY8IPItkNCP0iK8BqvmhSWV1pxpFqrB"
```

Now `usdc` can use your SHERPA/RoMEO API key! See the [getting started vignette](usdc.html) for usage of `usdc` functions.


### Setting up your API key in an `.Renviron` file

`.Renviron` file follow the same loading rules as `.Rprofile` files, the only difference is that it is is a file whose only purpose is to store environment variables. To use your API key in an `.Renviron` you have to type the following (note the absence of quotes):

```
BITQUERY_KEY=BQY8IPItkNCP0iK8BqvmhSWV1pxpFqrB
```

You can then check that your key has been found using `check_key()`:


```r
check_key()
#> [1] "BQY8IPItkNCP0iK8BqvmhSWV1pxpFqrB"
```
